Supporting Information for ``The dynamics of starvation and recovery''

### 2017-10-18 Manuscript/NatureComm/ReviewResponses.tex

\documentclass[ucm,12pt]{ucletter}

\setlength\parindent{1cm}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{xcolor}

%\name{Justin D. Yeakel}
\telephone{(209)~285-9571}
\email{jyeakel@ucmerced.edu}

\newcounter{section}
\newcounter{subsection}[section]
\setcounter{secnumdepth}{3}
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\Large\bfseries}}
\newcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\large\bfseries}}
\renewcommand \thesection{\@arabic\c@section}
\renewcommand\thesubsection{\thesection.\@arabic\c@subsection}

\makeatother

\begin{document}
\begin{letter}{
    Philosophical Transactions of the Royal Society B\\
    Publishing Section\\
    The Royal Society\\
    6-9 Carlton House Terrace\\
    London SW1Y 5AG\\
    England\\
    \centerline{\bf{Re: Eco-evolutionary dynamics and collective dispersal:}}\\
     \centerline{\bf{implications for salmon metapopulation robustness}}
}

\opening{To the editorial committee at Nature Communications,}

We thank both the Reviewers for their insightful comments and criticisms. Below, we address each one, point by point. We believe that the edits that we have made have increased the clarity of our contribution, in no small part due to the thought and detail exercised by the Reviewers. We hope that our responses satisfy the concerns listed below. Please do not hesitate to contact us if there are additional clarifications required.

\setstretch{1.2}

\section*{Reviewer I}
\subsection*{Major comments:}
\noindent \textcolor{cyan}{
{\bf Comment 1.1} stuff
}

\noindent {\bf Response 1.1} Stuff
\noindent To clarify, we have made the following edits to the text:\\
\noindent {\bf Line 111:} \emph{The partial derivative in Eq. 4 determines how the mean trait changes through time due to natural selection [ref], which is proportional to the change in mean fitness with respect to $\mu_i$.
We note that the derivative is dependent on both $\mu_i$ and $\mu_j$ due to the influence of the logarithm on the sum.}

\noindent \textcolor{cyan}{
{\bf Comment 1.2} Stuff
}

\noindent {\bf Response 1.2} Stuff

\vspace{5mm}

\singlespacing
\closing{Sincerely,\\
\fromsig{\includegraphics[scale=0.2]{signature.jpg}}\\
\fromname{
Justin D. Yeakel\\
Christopher P. Kempes\\
Sidney Redner}
}

\end{letter}
\end{document}

### 2017-10-18 NatureComm_REVISION/CoverLetter/YKR_cover.tex
new revision

\begin{itemize}
  \item Across the observed body size range, terrestrial mammals occupy a region of parameter space where sustained cyclic dynamics are not permitted. Relevant to observations of cycles in natural populations, the NSM predicts  that small mammal populations should be more prone to — whereas large mammals are buffered against — large transient oscillations, and this is a consequence of the timescales of starvation vs. recovery, both of which are governed by metabolism.
  \item Even transient cyclic dynamics increase the probability of extinction, and we argue that mammals should have energetic rates that avoid parameter regions close to transient oscillatory regimes. Moreover, regions of parameter space with small steady-state population sizes also increase the extinction probability. Thus we define a ``refuge'' of parameter space where the extinction probability is minimized between high probabilities of extinction via transience and high probabilities of extinction via low steady state densities, and show that mammals have allometrically constrained rates of starvation and recovery that place them squarely within this lower-risk space.
  \item Steady-state population sizes in the NSM support the energy equivalence hypothesis (equal energy use across populations of species with different individual body size) for a wide range of body sizes. Our model predicts that this energy equivalence breaks down at the smallest and largest body sizes observed for mammals, suggesting that there are strong dynamical constraints on mammalian body size. 
  \item We further investigate dynamical constraints on the range of mammalian body sizes by considering the invasion of a steady-state population by mammals with altered amounts of endogenous energetic reserves. We show that mammals with greater amounts of endogenous reserves are able to invade up to the size of the largest recorded mammal (ca. $10^7$g) above which leaner individuals can invade, such that this maximum body size bound serves as an evolutionary stable state. This suggests that the dynamics of starvation and recovery described in the NSM may provide a within-lineage mechanism for the evolution of larger body size among terrestrial mammals, a well-documented phenomenon known as Cope’s rule. This result is also supported by our finding that the steady state resource abundance is decreasing for larger mammals, which, in the context of classic resource competition theory, suggests that larger mammals will outcompete smaller ones given a shared resource. 
\end{itemize}

Our theory is analytic, relatively simple, and synthesizes a variety of fundamental ecological theories (energy equivalence hypothesis, resource competition theory, Cope’s rule, the fasting endurance hypothesis) and shows how many of these are the natural consequence of a simple dynamical framework that incorporates the energetic consequences of starvation and recovery combined with allometric timescales. We believe that our findings will be of general interest to ecologists, evolutionary biologists, and paleoecologists, such that consideration for publication in a broad interest journal such as PNAS is warranted.

@@

Suggested Referees:
\begin{itemize}
\item[] Alan Hastings, UC Davis, amhastings@ucdavis.edu
\item[] Felisa Smith, University of New Mexico, fasmith@unm.edu
\item[] Richard Sibly, University of Reading, r.m.sibly@reading.ac.uk
\item[] John DeLong, University of Nebraska, jpdelong@unl.edu
\item[] Seth Finnegan, Integrative Biology, UC Berkeley, sethf@berkeley.edu
\item[] Van Savage, University of California Los Angeles, vsavage@ucla.edu
\item[] Pablo Marquet, Department of Ecology, Pontificia Universidad Cat\`olica de Chile, pmarquet@bio.puc.cl
\end{itemize}

>>>>

\begin{itemize}
  \item Across the observed body size range, terrestrial mammals occupy a region of parameter space where sustained cyclic dynamics are not permitted.
  \item Even transient cyclic dynamics increase the probability of extinction, and we argue that mammals should have energetic rates that avoid parameter regions close to transient oscillatory regimes. Moreover, regions of parameter space with small steady-state population sizes also increase the extinction probability. Thus we define a ``refuge'' of parameter space where the extinction probability is minimized between high probabilities of extinction via transience and high probabilities of extinction via low steady state densities, and show that mammals have allometrically constrained rates of starvation and recovery that place them squarely within this lower-risk space. 
  \item We show that this lower-risk space becomes smaller for mammals with larger body size, thus providing a mechanistic explanation for the slightly higher extinction risk observed among larger mammals over evolutionary time.
  \item Steady-state population sizes in the NSM support the energy equivalence hypothesis (equal energy use across populations of species with different individual body size) for a wide range of body sizes. Our model predicts that this energy equivalence breaks down at the smallest and largest body sizes observed for mammals, suggesting that there are strong dynamical constraints on mammalian body size. 
  \item We further investigate dynamical constraints on the range of mammalian body sizes by considering the invasion of a steady-state population by mammals with altered amounts of endogenous energetic reserves. We show that mammals with greater amounts of endogenous reserves are able to invade up to the size of the largest recorded mammal (predicted $1.748\times10^7$g compared to an observed $1.74\times10^7$g estimated for \emph{Deinotherium}) above which leaner individuals can invade. This suggests that the dynamics of starvation and recovery described in the NSM may provide a within-lineage mechanism for the evolution of larger body size among terrestrial mammals, a well-documented phenomenon known as Cope's rule.
\end{itemize}

Our theory is analytic, relatively simple, and synthesizes a variety of fundamental ecological theories (energy equivalence hypothesis, resource competition theory, Cope's rule, the fasting endurance hypothesis) and shows how many of these are the natural consequence of a simple dynamical framework that incorporates the energetic consequences of starvation and recovery combined with allometric timescales. We believe that our findings will be of general interest to ecologists, evolutionary biologists, and paleoecologists, such that consideration for publication in a broad interest journal such as \emph{Nature Communications} is warranted.

@@

Suggested Referees:
\begin{itemize}
\item[] Felisa Smith, University of New Mexico, fasmith@unm.edu
\item[] Richard Sibly, University of Reading, r.m.sibly@reading.ac.uk
\item[] John DeLong, University of Nebraska, jpdelong@unl.edu
\item[] Seth Finnegan, Integrative Biology, UC Berkeley, sethf@berkeley.edu
\item[] Van Savage, University of California Los Angeles, vsavage@ucla.edu
\item[] Pablo Marquet, Department of Ecology, Pontificia Universidad Cat\`olica de Chile, pmarquet@bio.puc.cl
\end{itemize}
